\chapter{Introduction}
\label{intro} % Always give a unique label


Common ground. Teleoperated da Vinci surgical system is a robot-assisted surgical system that enhances surgeons performance in minimally invasive surgeries by allowing highly precise translation of surgeon`s hand movements to the instrument's movements. 

Practical problem condition. The currently available da Vinci surgery system has a laparoscopic camera, providing visual feedback to guide doctors during surgery. However, the system does not have any kinesthetic or cutaneous feedback, known as haptics.\cite{_intuitive_2018} 

The costs of that condition. 
Paraphrase
In manual minimally invasive surgery (MIS), surgeons feel the interaction of the instrument with the patient via a long shaft, which eliminates tactile cues and masks force cues. Some studies have linked the lack of significant haptic feedback in MIS to increased intraoperative injury. In teleoperated robot-assisted minimally invasive surgery (RMIS), all natural haptic feedback is eliminated because the surgeon no longer manipulates the instrument directly.\cite{okamura_haptic_2009} 

It is believed that the addition of haptic feedback in the da Vinci surgery robot will help to reduce the amount of surgical errors and intra-operative injuries, which will lead to faster post-surgery recovery time and decreased rate of unsuccessful surgeries. \cite{reiley_effects_2008, van_der_meijden_value_2009, okamura_haptic_2009}

Gist of your solution. There are many technical challenges to overcome in order to implement the haptic feedback in da Vinci robot. One of them is getting accurate force readings from the patient side manipulator (PSM). To address this issue, we are trying to create force-feedback device, that can be easily added to the existing surgery system.
